---
title: "DSCRTP_Pinnacles"
author: "Morgan Will"
date: "4/5/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

```{r}
#Packages used
library(vegan)
library(dummies)
library(maps)
library(sp)
library(maptools)
library(rgdal)
library(lattice)
library(classInt)
library(raster)
library(pscl)
library(MASS)
library(GGally)
library(dplyr)
library(easyCODA)
library(DescTools)
```

```{r}
```{r}
#Rasters of interest
DEPTH <- raster(paste0("../Data/USGS_bathy_raster/bathy.tif"))
Backscatter <- raster(paste0("../Data/USGS_mos_raster/mos.tif"))
Slope <- raster(paste0("../Data/USGS_slope_raster/slope.tif"))
Aspect <- raster(paste0("../Data/USGS_aspect_raster/aspect.tif"))

#Build raster stack

BACKSCATTER <- raster(vals=values(Backscatter), ext=extent(DEPTH),
                      crs=crs(DEPTH), nrows=dim(DEPTH)[1], 
                      ncols=dim(DEPTH)[2])
SLOPE <- raster(vals=values(Slope), ext=extent(DEPTH),
                crs=crs(DEPTH), nrows=dim(DEPTH)[1], 
                ncols=dim(DEPTH)[2])
ASPECT <- raster(vals=values(Aspect), ext=extent(DEPTH),
                 crs=crs(DEPTH), nrows=dim(DEPTH)[1],
                 ncols=dim(DEPTH)[2])

# now replacing zeros with NA
BACKSCATTER <- calc(BACKSCATTER, function(x) ifelse(x == 0, NA, x))

# now cover the other layers with NA's based upon DEPTH
DEPTH <- mask(DEPTH, BACKSCATTER)
SLOPE <- mask(SLOPE, BACKSCATTER)
ASPECT <- mask(ASPECT, BACKSCATTER)

AllRasters <- stack(DEPTH, BACKSCATTER, SLOPE, ASPECT)
names(AllRasters) <- c('DEPTH', 'BACKSCATTER', 'SLOPE', 'ASPECT')
View(AllRasters)
```

```{r}
#DSCRTP data clipped to Pinnacles region and filtered by genus in ArcGIS. Imported here:
Muricea <- read.csv("../Data/Pinnacles_Muricea_Final.csv")
Muricea  <- SpatialPointsDataFrame(Muricea [ , c("longitude__degrees_east_", "latitude__degrees_north_")],
                              Muricea,
                              proj4string = CRS(proj4string(AllRasters)))
View(Muricea)
Swiftia <- read.csv("../Data/Pinnacles_Swiftia_Final.csv")
Swiftia  <- SpatialPointsDataFrame(Swiftia [ , c("longitude__degrees_east_", "latitude__degrees_north_")],
                              Swiftia,
                              proj4string = CRS(proj4string(AllRasters)))
View(Swiftia)
Thesea <- read.csv("../Data/Pinnacles_Thesea_Final.csv")
Thesea  <- SpatialPointsDataFrame(Thesea [ , c("longitude__degrees_east_", "latitude__degrees_north_")],
                              Thesea,
                              proj4string = CRS(proj4string(AllRasters)))
View(Thesea)
```





##Visualizing the Data

```{r}
#Depth
ggplot(AllGroups, aes(x=DepthInMeters__m_))+
  geom_histogram(color="darkblue", fill="lightblue")+
  facet_grid(Genus~.)
#Backscatter
ggplot(AllGroups, aes(x=BACKSCATTER))+
  geom_histogram(color="darkblue", fill="lightblue")+
  facet_grid(Genus~.)
#Slope
ggplot(AllGroups, aes(x=SLOPE))+
  geom_histogram(color="darkblue", fill="lightblue")+
  facet_grid(Genus~.)
#Aspect
ggplot(AllGroups, aes(x=ASPECT))+
  geom_histogram(color="darkblue", fill="lightblue")+
  facet_grid(Genus~.)
#SST --> very coarse data set, need to find a better source
ggplot(AllGroups, aes(x=SST))+
  geom_histogram(color="darkblue", fill="lightblue")+
  facet_grid(Genus~.)
#Observation Year
ggplot(AllGroups, aes(x=ObservationYear))+
  geom_histogram(color="darkblue", fill="lightblue")+
  facet_grid(Genus~.)
```

##Modeling

```{r}
#Muricea
plot(Muricea$latitude__degrees_north_,Muricea$longitude__degrees_east_)
#Form a data frame of muricea counts (OID_=site, IndividualCount=abundance) and envirmental data
Muricea_raw <- data.frame(Muricea$OID_, Muricea$IndividualCount)
Muricea_env <- data.frame(Muricea$OID_, Muricea$DEPTH, Muricea$BACKSCATTER, Muricea$SLOPE, Muricea$ASPECT, Muricea$SST, Muricea$ObservationYear)
#Compare with rda
Muricea_rda <- rda(Muricea_raw, Muricea_env)
plot(Muricea_rda)
anova(Muricea_rda)

# extract enviornment for model building
muricea_env <- as.data.frame(extract(AllRasters, Muricea))

#General Linear Model (Gaussian)
Muricea_gaussian <- glm(Muricea$IndividualCount ~  DEPTH + BACKSCATTER + SLOPE + ASPECT ,
                        data=muricea_env, family=gaussian)
summary(Muricea_gaussian)

#observed predicted plots
plot(Muricea$IndividualCount, predict(Muricea_gaussian))



#General Linear Model (Poisson)
Muricea_poisson <- glm(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=Muricea, family=poisson)
summary(Muricea_poisson)
#General Linear Model (Negative Binomial)
Muricea_nb <- glm.nb(IndividualCount ~ DEPTH + BACKSCATTER + SLOPE + ASPECT, data=Muricea)
summary(Muricea_nb)
#Compare Models
AIC(Muricea_gaussian, Muricea_poisson, Muricea_nb) #nb had lowest AIC, so best fit (penalizes for model complexity)
PseudoR2(Muricea_gaussian, which = NULL) #0.02077
PseudoR2(Muricea_poisson, which = NULL) #0.13157 --> best fit based on pseudo r2
PseudoR2(Muricea_nb, which = NULL) #0.04123
#In solutions, predicted vs observed

#Generate raster
Muricea_raster <- as.data.frame(AllRasters)
View(as.data.frame(Muricea_raster))
tmp <- predict(AllRasters, Muricea_nb)
```


```{r}
#Swiftia
plot(Swiftia$latitude__degrees_north_,Swiftia$longitude__degrees_east_)
#Form a data frame of swiftia counts (OID_=site, IndividualCount=abundance) and envirmental data
Swiftia_raw <- data.frame(Swiftia$OID_, Swiftia$IndividualCount)
Swiftia_env <- data.frame(Swiftia$DEPTH, Swiftia$BACKSCATTER, Swiftia$SLOPE, Swiftia$ASPECT, Swiftia$SST, Swiftia$ObservationYear)
#Compare with rda
Swiftia_rda <- rda(Swiftia_raw, Swiftia_env)
anova(Swiftia_rda)
plot(Swiftia_rda) #this plot is hard to read, trying to make a fancier one below

#Compare with PCA
Swiftia_pca <- prcomp(Swiftia[,c(1,55,117,118,119,120,121)], center=TRUE, scale.=TRUE)
summary(Swiftia_pca)

#General Linear Model (Gaussian)
Swiftia_gaussian <- glm(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=Swiftia, family=gaussian)
summary(Swiftia_gaussian)
#General Linear Model (Poisson)
Swiftia_poisson <- glm(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=Swiftia, family=poisson)
summary(Swiftia_poisson)
#General Linear Model (Negative Binomial)
Swiftia_nb <- glm.nb(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=Swiftia)
summary(Swiftia_nb)
#Compare Models
AIC(Swiftia_gaussian, Swiftia_poisson, Swiftia_nb) #nb had lowest AIC, so best fit
PseudoR2(Swiftia_gaussian, which = NULL) #0.00874
PseudoR2(Swiftia_poisson, which = NULL) #0.09613 --> best fit based on pseudo r2
PseudoR2(Swiftia_nb, which = NULL) #0.04355
```

```{r}
#Thesea
plot(Thesea$latitude__degrees_north_, Thesea$longitude__degrees_east_)
#Form a data frame of thesea counts (OID_=site, IndividualCount=abundance) and envirmental data
Thesea_raw <- data.frame(Thesea$OID_, Thesea$IndividualCount)
Thesea_env <- data.frame(Thesea$DEPTH, Thesea$BACKSCATTER, Thesea$SLOPE, Thesea$ASPECT, Thesea$SST, Thesea$ObservationYear)
#Compare with rda
Thesea_rda <- rda(Thesea_raw, Thesea_env)
plot(Thesea_rda)
anova(Thesea_rda)

#General Linear Model (Gaussian)
Thesea_gaussian <- glm(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=Thesea, family=gaussian)
summary(Thesea_gaussian)
#General Linear Model (Poisson)
Thesea_poisson <- glm(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=Thesea, family=poisson)
summary(Thesea_poisson)
#General Linear Model (Negative Binomial)
Thesea_nb <- glm.nb(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=Thesea)
summary(Thesea_nb)
#Compare Models
AIC(Thesea_gaussian, Thesea_poisson, Thesea_nb) #nb had lowest AIC, so best fit
PseudoR2(Thesea_gaussian, which = NULL) #0.02670
PseudoR2(Thesea_poisson, which = NULL) #0.10877 --> best fit based on pseudo r2
PseudoR2(Thesea_nb, which = NULL) #0.06573
```
```{r}
#Evaluating all three groups in a single model
AllGroups <- rbind(Muricea, Swiftia, Thesea)
#Form a data frame of counts (OID_=site, IndividualCount=abundance) and environmental data
AllGroups_raw <- data.frame(AllGroups$OID_, AllGroups$IndividualCount)
AllGroups_env <- data.frame(AllGroups$DEPTH, AllGroups$BACKSCATTER, AllGroups$SLOPE, AllGroups$ASPECT, AllGroups$SST, AllGroups$ObservationYear)
#Compare with rda
AllGroups_rda <- rda(AllGroups_raw, AllGroups_env)
plot(AllGroups_rda)
anova(AllGroups_rda)

#General Linear Model (Gaussian)
AllGroups_gaussian <- glm(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=AllGroups, family=gaussian)
summary(AllGroups_gaussian)
#General Linear Model (Poisson)
AllGroups_poisson <- glm(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=AllGroups, family=poisson)
summary(AllGroups_poisson)
#General Linear Model (Negative Binomial)
AllGroups_nb <- glm.nb(IndividualCount ~ OID_ + DEPTH + BACKSCATTER + SLOPE + ASPECT + SST + ObservationYear, data=AllGroups)
summary(AllGroups_nb)
#Compare Models
AIC(AllGroups_gaussian, AllGroups_poisson, AllGroups_nb) #nb had lowest AIC, so best fit
PseudoR2(AllGroups_gaussian, which = NULL) #0.00501
PseudoR2(AllGroups_poisson, which = NULL) #0.05437 --> best fit based on pseudo r2
PseudoR2(AllGroups_nb, which = NULL) #0.02665
```

##Compare Single Group vs Multi-Group Models

```{r}
#AIC (smaller value = better fit)
AIC(Muricea_nb, Swiftia_nb, Thesea_nb, AllGroups_nb)
```

