---
title: "DSCRTP_Filtering"
author: "Morgan Will"
date: "3/24/2022"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)
opts_knit$set(root.dir='../')
```

```{r}
#Packages Used
library(plyr)    #Version 1.8.6
library(readr)   #Version 2.1.2
library(ggplot2) #Version 3.3.5
library(GGally)  #Version 2.1.2
library(dplyr)   #Version 1.0.8
library(mlbench) #Version 2.1.3
```


```{r}
#Set working directory to local project location
setwd("/Users/morganwill/Desktop/Pinnacles_Trend")
#Inport raw data from DSCRTP (Note: "DSCRTP_AllData" is not included in the Git Repo due to its large size. To download this data, go to https://www.ncei.noaa.gov/maps/deep-sea-corals/mapSites.htm)
DSCRTP_AllData <- read.csv("../Data/DSCRTP_AllData.csv", header = TRUE)
```

```{r}
#First row indicates units, need to remove for further analysis:
DSCRTP_AllData <- DSCRTP_AllData[2:nrow(DSCRTP_AllData),]
head(DSCRTP_AllData)
```

# Current data is world-wide, filter to Gulf of Mexico (GOM)
* Potential columns to use: "Ocean", "LargeMarineEcosystem", "Country", "FishCouncilRegion", Locality", Longitude", "Latitude"
* "FishCouncilRegion" determined to be the best option, due to relatively coarse gradient and percent of entries with this field filled (96.66%)
```{r}
#Filter data based on FishCouncilRegion --> FishCouncilRegion = "Gulf of Mexico"
DSCRTP_GOM <- DSCRTP_AllData %>% filter(FishCouncilRegion=="Gulf of Mexico")
glimpse(DSCRTP_GOM)
#678,868 rows in AllData --> 32,991 rows in GOM
#Quick check that filter worked properly using map:

```

# Filter data for depth range of interest 
--> Using genus level IDs, mainly because species level IDs of these species are a bit rough. However, the depth ranges that these species live in can be used to filter the genus level data a bit further (ex: I am interested in Swiftia exserta, a mesophotic species, so can filter out the deep-sea swiftia species using depth range identifiers)
* Potnetial columns to use: "DepthInMeters", "MinimumDepthInMeters", "MaximumDepthInMeters"
* 
```{r}

```

# Filter Data to species of interest 
--> Going to stay at genus level: Swiftia, Muricea, Thesea, and Paramuricea
Potential columns to use:
* ScientificName
* VerbatimScientificName
* Genus
* Species
```{r}

```

# Now that the data is filtered down a bit, explore what is left.
Columns used for filtering:
* 
Columns to use moving forward:
* ObservationDate
* ObservationYear
* ObservationTime
* Vessel
* PI
* PIAffiliation
* Sampling Equipment
* SampleAreaInSquareMeters
* footprintWKT
* footprintSRS
* IndividualCount
* CategoricalAbundance
* Density
* Cover
* Condition
* StartLatitude
* StartLongitude
* EndLatitude
* EndLongitude
* VerbatimLatitude
* VerbatimLongitude
* LocationAccuracy
* Temperature
* Salinity
* Oxygen
* pH
* pCO2
* TA
* DIC
* RecordType
```{r}

```



