---
title: "DSCRTP_Filtering"
author: "Morgan Will"
date: "3/24/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Set working directory to local project location
setwd("/Users/morganwill/Desktop/Pinnacles_Area")
#Inport raw data from DSCRTP (Note: "DSCRTP_AllData" is not included in the Git Repo due to its large size. To download this data, go to https://www.ncei.noaa.gov/maps/deep-sea-corals/mapSites.htm)
DSCRTP_AllData <- read.csv("Data/DSCRTP_AllData.csv", header = TRUE)
```

```{r}
#First row indicates units, need to remove for further analysis:
DSCRTP_AllData <- DSCRTP_AllData[2:nrow(DSCRTP_AllData),]
head(DSCRTP_AllData)
```
# Current data is world-wide, filter to Gulf of Mexico (GOM)
Potential columns to use:
* Ocean
* LargeMarineEcosystem
* Country
* FishCouncilRegion
* Locality
* Longitude
* Latitude
```{r}

```
# Filter Data to species of interest 
--> Going to stay at genus level: Swiftia, Muricea, Thesea, and Paramuricea
Potential columns to use:
* ScientificName
* VerbatimScientificName
* Genus
* Species
```{r}

```
# Filter data for depth range of interest 
--> Using genus level IDs, mainly because species level IDs of these species are a bit rough. However, the depth ranges that these species live in can be used to filter the genus level data a bit further (ex: I am interested in Paramuricea biscaya, a deep sea species, so can filter out the mesophotic paramuricea species using depth range identifiers)
Potnetial columns to use:
* DepthInMeters
* MinimumDepthInMeters
* MaximumDepthInMeters
```{r}

```
# Now that the data is filtered down a bit, explore what is left.
Columns used for filtering:
* 
Columns to use moving forward:
* ObservationDate
* ObservationYear
* ObservationTime
* Vessel
* PI
* PIAffiliation
* Sampling Equipment
* SampleAreaInSquareMeters
* footprintWKT
* footprintSRS
* IndividualCount
* CategoricalAbundance
* Density
* Cover
* Condition
* StartLatitude
* StartLongitude
* EndLatitude
* EndLongitude
* VerbatimLatitude
* VerbatimLongitude
* LocationAccuracy
* Temperature
* Salinity
* Oxygen
* pH
* pCO2
* TA
* DIC
* RecordType
```{r}

```



